---
title: "QQ Midterm Test"
author: "Anthony Mensah"
date: "2024-01-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MASS)
library(lsmeans)
library(dplyr)
library(car)
```

## Q1

```{r echo=FALSE}
# Create table for isolated_island

iso <- data.frame(
  trichromacy = c(0.35, 0.53),
  color_deficiency = c(0.1, 0.02)
)

# set row_names 
row.names(iso) <- c("male", "female")

print(iso)
```

```{r include=FALSE}
total_tri <- sum(iso$trichromacy)
total_coldef <- sum(iso$color_deficiency)
total_mals <- rowSums(iso[row.names(iso) == "male", c("trichromacy", "color_deficiency")])
total_fems <- rowSums(iso[row.names(iso) == "female", c("trichromacy", "color_deficiency")])

```

## 1a.

```{r echo=FALSE}
# P(color_deficiency)
p_color_deficiency <- total_coldef/(sum(total_coldef, total_tri))
print(paste("P(color_deficiency) =", p_color_deficiency))

# P(trichromacy)
p_trichromacy <- total_tri/(sum(total_coldef, total_tri))
print(paste("P(trichromacy) =", p_trichromacy))

# P(male)
p_male <- total_mals/(sum(total_mals, total_fems))
print(paste("P(male) =", p_male))

# P(female)
p_female <- total_fems/(sum(total_mals, total_fems))
print(paste("P(female): =", p_female))

# P(color_deficiency, Male)
p_cdm <- p_color_deficiency * p_male
print(paste("P(color_deficiency, Male) =", p_cdm))

# P(color_deficiency, female)
p_cdf <- p_color_deficiency * p_female
print(paste("P(color_deficiency, female) =", p_cdf))

# P(color_deficiency | male)
p_cd_m <- p_cdm/p_male
print(paste("P(color_deficiency | male) =", p_cd_m))

# P(color_deficiency | female)
p_cd_f <- p_cdf/p_female
print(paste("P(color_deficiency | female) =", p_cd_f))

```

## 1b.

Is  P(color_deficiency, Male) equal to P(color_deficiency) * P(male)? 
Yes. computationally, P(color_deficiency, Male) equals 0.054 likewise, P(color_deficiency) * P(male) equals 0.054

Is P(color_deficiency, Male) equal to  P(color_deficiency | male) * P(male)?
Yes. P(color_deficiency, Male) 0.054 is the same as P(color_deficiency | male) * P(male) 0.054

## 1c.

Based on these computations, the probability of color deficiency is independent on the gender category because with regards to each gender (male, female), the probability of color deficiency is the same.
that is 0.12


## Q2

```{r echo=FALSE}
data_hand <- data.frame(
  right_handed = c(0.36, 0.44),
  left_handed = c(0.09, 0.11)
)

row.names(data_hand) <- c("male", "female")
print(data_hand)

```

```{r include=FALSE}
t_rh <- sum(data_hand$right_handed)
t_lh <- sum(data_hand$left_handed)
t_mrl <- rowSums(data_hand[row.names(data_hand) == "male", c("right_handed", "left_handed")])
t_frl <- rowSums(data_hand[row.names(data_hand) == "female", c("right_handed", "left_handed")])

```

## 2a.

```{r echo=FALSE}
# P(left_handed) = 
p_lh <- t_lh/sum(t_lh, t_rh)
print(paste("P(left_handed) =",p_lh))

# P(right_handed) = 
p_rh <- t_rh/sum(t_rh, t_lh)
print(paste(" P(right_handed) =",p_rh))

# P(male)
p_m <- t_mrl/sum(t_mrl, t_frl)

# p(female)
p_f <- t_frl/sum(t_mrl, t_frl)

# P(left_handed, male) = 
p_lhm <- p_lh * p_m
print(paste("P(left_handed, male) =",p_lhm))

# P(right_handed, female) = 
p_rhf <- p_rh * p_f 
print(paste("P(right_handed, female) =",p_rhf ))

# P(left_handed | male) = 
p_lh_m <- (p_lhm / p_m)
print(paste("P(left_handed | male) =",p_lh_m ))

# P(right_handed | female) =
p_rh_f <- p_rhf/p_f
print(paste("P(right_handed | female) =",p_rh_f ))

```

## 2b.

Is P(left_handed, male) equal to P(left_handed) \* P(male) ?
Yes.
Statistically, P(left_handed, male) is given as P(left_handed) \* P(male) which in turn equals 0.09

Is P(left_handed, male) equal to P(left_handed \| male) \* P(male)?
Yes.
Statistically, P(left_handed \| male) \* P(male) is reduced to P(left_handed, male) which is 0.09

## 2c.

From the conditional probabilities, we can say that the probability of being left-handed or right-handed is independent of gender because P(left_handed \| male) = P(left_handed) and P(right_handed \| female) = P(right_handed).
In other words, knowing the gender does not provide additional information about handedness.
If handedness and gender were dependent, the conditional probabilities would not be equal to the marginal probabilities of handedness.

## Q3
Behavioral experiment with 3 choices: Probability of correct response is 1/3. Type of incorrect response is irrelevant

## 3a.

```{r echo=FALSE}
# Probability of 2 correct responses out of 6 trials
# Set parameters. let
n <- 6  # total number of trials
k <- 2  # number of successes (correct responses)
probability_of_success <- 1/3  # probability of success (correct response)

# Calculate the probability
probability <- dbinom(k, size = n, prob = probability_of_success)

# Print the result
print(paste("Probability of 2 correct responses out of 6 trials =", fractions(probability)))

```

## 3b.

```{r echo=FALSE}
# Probability of 2 correct responses out of 6 trials
# Set parameters. let
n <- 6  # total number of trials
k <- 3  # number of successes (correct responses)
probability_of_success <- 1/3  # probability of success (correct response)

# Calculate the probability
probability <- dbinom(k, size = n, prob = probability_of_success)

# Print the result
print(paste("Probability of 3 correct responses out of 6 trials =", fractions(probability)))
```

## 3c.

```{r echo=FALSE}
# Probability of 20 correct responses out of 60 trials
# Set parameters. let
n <- 60  # total number of trials
k <- 20  # number of successes (correct responses)
probability_of_success <- 1/3  # probability of success (correct response)

# Calculate the probability
probability <- dbinom(k, size = n, prob = probability_of_success)

# Print the result
print(paste("Probability of 20 correct responses out of 60 trials =", fractions(probability)))
```

## 3d.

```{r echo=FALSE}
# Probability of 20 correct responses out of 60 trials
# Set parameters. let
n <- 60  # total number of trials
k <- 30  # number of successes (correct responses)
probability_of_success <- 1/3  # probability of success (correct response)

# Calculate the probability
probability <- dbinom(k, size = n, prob = probability_of_success)

# Print the result
print(paste("Probability of 30 correct responses out of 60 trials =", fractions(probability)))
```

## Q4

```{r include=FALSE}
dTmQ4 = read.delim("C:/Users/antho/Downloads/QQ Data/dataTestMidQ4.txt", header=TRUE, sep="\t")

dTmQ4$Measure
```

## 4a.

```{r echo=FALSE}
cat("The snippet length(dTmQ4$Measure)")
cat("Returns count of the ROWS of the Measure column in dTmQ4:", length(dTmQ4$Measure))

```

## 4b.

```{r echo=FALSE}
cat("Variance =",var(dTmQ4$Measure), "\n")
cat("Standard Deviation =", sd(dTmQ4$Measure), "\n")
cat("Mean =", mean(dTmQ4$Measure), "\n")
cat("Median =", median(dTmQ4$Measure), "\n")

q <- quantile(dTmQ4$Measure, probs = c(0, 0.25, 0.5, 0.75, 1))
# Compute quartile ranks
quartile_ranks <- cut(dTmQ4$Measure, breaks = q, labels = FALSE, include.lowest = TRUE)
cat("First Quartile (Q1) =", q[2], "\n")
cat("Third Quartile (Q3) =", q[4], "\n")

```

## 4c.

```{r echo=FALSE}
cat("Standard Error =", sd(dTmQ4$Measure) / sqrt(length(dTmQ4$Measure)), "\n")

# Perform t-test and calculate 95% CI
t_test_result_90 <- t.test(dTmQ4$Measure, conf.level = 0.90)

# Extract the confidence interval from the result
confidence_interval_90 <- t_test_result_90$conf.int

# Print the result
cat("90% Confidence Interval:", confidence_interval_90, "\n")


# Perform t-test and calculate 95% CI
t_test_result <- t.test(dTmQ4$Measure)

# Extract the confidence interval from the result
confidence_interval <- t_test_result$conf.int

# Print the result
cat("95% Confidence Interval:", confidence_interval, "\n")

```

## Q5

The 3 two-sample unpaired t-tests (5a, 5b, 5c) comparing 2 conditions

```{r echo=FALSE}
# Load data
dTmQ5 = read.delim("C:/Users/antho/Downloads/QQ Data/dataTestMidQ5.txt", header=TRUE, sep="\t")

dTmQ5Cond1 = dTmQ5$Resp[dTmQ5$Conditions == "Cond1"]
dTmQ5Cond2 = dTmQ5$Resp[dTmQ5$Conditions == "Cond2"]

ttest_5a <- t.test(dTmQ5Cond1, dTmQ5Cond2, paired = FALSE, var.equal = TRUE)
ttest_5b <- t.test(dTmQ5Cond1, dTmQ5Cond2, paired = FALSE, var.equal = FALSE)
ttest_5c <- t.test(dTmQ5Cond1, dTmQ5Cond2, paired = FALSE)

cat("5a is a student's t-test as a result of pre defining variance as EQUAL", "\n")
cat("5b is Welch's t-test with variance pre defined as NOT EQUAL","\n")
cat("5c is also Welch's t-test, though equality of variance was not defined prior in", "\n", "the function, but the default value for unpaired t-test variance is always NOT EQUAL","\n")
```

## Q6
## 6a.
No. student S does not come to class due ONLY to the bad weather

## 6b.
No. Student S comes to class only when the condition is TRUE

## 6c.
Inconclusive(Yes and No): Student S could miss class for other reasons since the statement does not explicitly say student S misses class strictly for ONLY bad weather.

## 6d.
No, student S does not come to class if the weather is not fine